---
title: ""
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=15, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(lubridate)
library(pracma)
library(patchwork)
library(RColorBrewer)
library(ggpubr)

cations <- read_csv("anion_cation/cation_xlsx.csv") %>% 
  mutate(
    MeasurementDate = mdy(MeasurementDate), 
    Date = mdy(Date),
    Calcium = as.numeric(Calcium), Magnesium = as.numeric(Magnesium), Potassium = as.numeric(Potassium), Sodium = as.numeric(Sodium), Date = as.Date(Date))

anions <- read_csv("anion_cation/anion_xlsx.csv") %>% 
  mutate(
    MeasurementDate = mdy(MeasurementDate),
    Date_full = mdy(paste(Month, Day, Year, sep = " ")),
    Nitrate = as.numeric(Nitrate), Flouride = as.numeric(Flouride), Chloride = as.numeric(Chloride))

manual_well_creek <- read_csv("manual_to_dl.csv") %>% 
    mutate(Site = case_when(
    Well == "P1" ~ "1",
    Well == "P2" ~ "2",
    Well == "P3" ~ "3",
    Well == "P4" ~ "4",
    Well == "P5" ~ "5",
    Well == "C1" ~ "1",
    Well == "C2" ~ "2",
    Well == "C3" ~ "3",
    Well == "C4" ~ "4",
    Well == "C5" ~ "5"
  ),
  Area = case_when(
    str_sub(Well, 1, 1) == "P" ~ "Well",
    str_sub(Well, 1, 1) == "C" ~ "Creek"
  ), Date = mdy_hms(Date)) %>% 
  filter(SPC_before < 1000, year(Date) > 2024)

manual_before_after <- read_csv("manual_before_after.csv") %>% 
  mutate(Date = mdy_hms(Date))
  # filter(year(Date) > 2025)

dfs = list()
hobos_all = c("Dwnstrm", "P1", "P2", "P3", "Mdstrm", "Spring", "P4", "P5", "P5_creek_cond") 
for (site in hobos_all) {
  data = read_csv(paste0("hobos/hobos_running/", site, "_current.csv"))
  data = data %>% drop_na()
  to_filter = data$'Specific Conductivity'
  result = hampel(to_filter, k = 100, t0 = 2)
  final = data %>% slice(-result$ind)
  
  dfs[[site]] = final[, c("Date-Time", "Specific Conductivity", "Temperature")]
}
df_full = bind_rows(dfs, .id = "Site") %>% drop_na()
rm(list = c("dfs", "result", "final", "to_filter", "data"))
```


## Plot Set 1: Anions, Before Pump vs. Creek for P1, P3, & P5

```{r, fig.width=13}
build_plot = function(df, site, metric) {
  pl_data <- df %>% 
    filter(PumpTest_Creek != "A", Well == site) %>% 
    mutate(PumpTest_Creek = ifelse(PumpTest_Creek == "B", "Well", "Creek")) %>% 
    group_by(Date_full, PumpTest_Creek) %>% 
    summarize(avg = mean(.data[[metric]], na.rm = TRUE), .groups = "drop")
    
  l_lim <- min(df[[metric]], na.rm = TRUE)   
  u_lim <- max(df[[metric]], na.rm = TRUE)
  
  plot <- ggplot(pl_data, aes(x = Date_full, y = avg, fill = PumpTest_Creek)) + 
      geom_path(aes(color = PumpTest_Creek)) + 
      geom_point(pch = 21, size = 2.5) + 
      theme_bw(base_size = 16) + 
      theme(plot.subtitle = element_text(size = 10)) +
      guides(color = "none") +
      scale_y_continuous(limits = c(l_lim, u_lim)) +
      scale_x_date(limits = range(anions$Date_full, na.rm = TRUE), date_breaks = "4 months") +
      labs(title = paste0(metric, " at ", site, ": well vs. Creek"),
           subtitle = "Well measurement taken before pumping",
           fill = "Location",
           x = "Date",
           y = "Concentration (ppm)")
  
  return(plot)
}

p1_n = build_plot(anions, "P1", "Nitrate")
p2_n = build_plot(anions, "P3", "Nitrate")
p3_n = build_plot(anions, "P5", "Nitrate")

p1_s = build_plot(anions, "P1", "Sulfate")
p2_s = build_plot(anions, "P3", "Sulfate")
p3_s = build_plot(anions, "P5", "Sulfate")

p1_c = build_plot(anions, "P1", "Chloride")
p2_c = build_plot(anions, "P3", "Chloride")
p3_c = build_plot(anions, "P5", "Chloride")

wrap_plots(
  p1_n, p2_n, p3_n,
  p1_c, p2_c, p3_c,
  p1_s, p2_s, p3_s,
  nrow = 3,
  ncol = 3,
  guides = "collect",
  axes = "collect"
)

```

Nitrates in the P1 well are meaningfully higher than P3 and P5; additionally, P1 is the only instance where nitrates are higher in the well than the creek. Similarity in values and behavior suggests that Nitrate and Chloride concentrations are related, possibly due to their presence in fertilizer or application at similar times of year. Sulfate more or less operates in its own universe; overall concentrations tend to be slightly higher, and creek/well relationships flip except for at P1. 


## 2: SPC with anion overlays

```{r, fig.width=13}
double_graph = function(site){
anion_data = anions %>% 
  filter(PumpTest_Creek == "B", Well == site) %>% 
    summarize(avg_n = mean(Nitrate), avg_s = mean(Sulfate), avg_c = mean(Chloride), .by = Date_full)
spc_data = df_full %>% 
  filter(Site == site, `Date-Time` >= min(anion_data$Date_full), `Date-Time` <= max(anion_data$Date_full))

p1 = ggplot(spc_data, aes(x = `Date-Time`, y = `Specific Conductivity`)) + 
  geom_line() + 
  theme_bw(base_size = 20) +
  labs(title = paste0("SPC and anion data at ", site),
       x = NULL,
       y = "SPC (usm/cm)") +
    theme(plot.margin = margin(0,5,-5,5),
        plot.title.position = "plot")

p2 = ggplot(anion_data) + 
  geom_line(aes(x = Date_full, y = avg_n, color = "Nitrate")) + 
  geom_point(aes(x = Date_full, y = avg_n, fill = "Nitrate"), pch = 21, size = 2.5) + 
  geom_line(aes(x = Date_full, y = avg_s, color = "Sulfate")) + 
  geom_point(aes(x = Date_full, y = avg_s, fill = "Sulfate"), pch = 21, size = 2.5) + 
  geom_line(aes(x = Date_full, y = avg_c, color = "Chloride")) + 
  geom_point(aes(x = Date_full, y = avg_c, fill = "Chloride"), pch = 21, size = 2.5) + 
  theme_bw(base_size = 20) + 
  guides(color = "none") +
  theme(plot.margin = margin(-5,5,0,5), 
        plot.title.position = "plot") +
  labs(title = NULL,
      x = "Date", 
      y = "Concentration (ppm)",
      fill = "Anion")
p1 / p2 + plot_layout(guides = "collect")
}

double_graph("P1")  
double_graph("P3") 
double_graph("P5") 
```

## Plot set 3: all cations

```{r, fig.width=13}
cat_plot = function(df, site, metric) {
  plot = df %>% 
    filter(PumpTest_Creek != "A", Well == site, Dilution == "1:4") %>% 
    mutate(PumpTest_Creek = ifelse(PumpTest_Creek == "B", "Well", "Creek"), Date = as.Date(Date)) %>% 
    group_by(Date, PumpTest_Creek) %>% 
    summarize(avg = mean(.data[[metric]], na.rm = TRUE), .groups = "drop") %>% 
    
  l_lim <- min(df[[metric]], na.rm = TRUE)   
  u_lim <- max(df[[metric]], na.rm = TRUE)
    
   ggplot(aes(x = Date, y = avg, fill = PumpTest_Creek)) + 
      geom_line(aes(color = PumpTest_Creek)) + 
      geom_point(pch = 21, size = 2.5) + 
      theme_bw(base_size = 16) + 
      theme(plot.subtitle = element_text(size = 10)) +
      scale_y_continuous(limits = c(l_lim, u_lim)) +
      scale_x_date(
        limits = c(min(cations$Date), max(cations$Date)),
            date_breaks = "1 month",
            date_labels = "%Y-%m") +
      labs(title = paste0(metric, " at ", site, ": well vs. Creek"),
           subtitle = "Well measurement taken before pumping",
           fill = "Location",
           x = "Date",
           y = "Concentration (ppm)") +
    guides(color = 'none')

  return(plot)
}

p1_p = cat_plot(cations, "P1", "Potassium")
p2_p = cat_plot(cations, "P3", "Potassium")
p3_p = cat_plot(cations, "P5", "Potassium")

p1_m = cat_plot(cations, "P1", "Magnesium")
p2_m = cat_plot(cations, "P3", "Magnesium")
p3_m = cat_plot(cations, "P5", "Magnesium")
 
p1_c = cat_plot(cations, "P1", "Calcium")
p2_c = cat_plot(cations, "P3", "Calcium")
p3_c = cat_plot(cations, "P5", "Calcium")

p1_s = cat_plot(cations, "P1", "Sodium")
p2_s = cat_plot(cations, "P3", "Sodium")
p3_s = cat_plot(cations, "P5", "Sodium")

wrap_plots(
  # p1_p, p2_p, p3_p,
  p1_m, p2_m, p3_m,
  p1_s, p2_s, p3_s,
  p1_c, p2_c, p3_c,
  nrow = 3,
  ncol = 3,
  guides = "collect",
  axes = "collect"
)

```

## Plot Set 4: Cations & SPC

```{r, fig.width=13}
plot_cations = function(site){
cation_data = cations %>% 
  filter(PumpTest_Creek == "B", Well == site) %>% 
    summarize(avg_c = mean(Calcium), avg_p = mean(Potassium), avg_m = mean(Magnesium), .by = Date)
spc_data = df_full %>% 
  filter(Site == site, `Date-Time` >= min(cation_data$Date), `Date-Time` <= max(cation_data$Date))

p1 = ggplot(spc_data, aes(x = `Date-Time`, y = `Specific Conductivity`)) + 
  geom_line() + 
  theme_bw(base_size = 20) +
  labs(title = paste0("SPC and cation data at ", site),
       x = NULL,
       y = "SPC (usm/cm)") +
    theme(plot.margin = margin(0,5,-5,5),
        plot.title.position = "plot")

p2 = ggplot(cation_data) + 
  geom_line(aes(x = Date, y = avg_c, color = "Calcium")) + 
  geom_point(aes(x = Date, y = avg_c, fill = "Calcium"), pch = 21, size = 2.5) + 
  # geom_line(aes(x = Date, y = avg_p, color = "Potassium")) + 
  # geom_point(aes(x = Date, y = avg_p, fill = "Potassium"), pch = 21, size = 2.5) + 
  geom_line(aes(x = Date, y = avg_m, color = "Magnesium")) + 
  geom_point(aes(x = Date, y = avg_m, fill = "Magnesium"), pch = 21, size = 2.5) + 
  theme_bw(base_size = 20) + 
  guides(color = "none") +
  theme(plot.margin = margin(-5,5,0,5), 
        plot.title.position = "plot") +
  labs(title = NULL,
      x = "Date", 
      y = "Concentration (ppm)",
      fill = "Cation")
p1 / p2 + plot_layout(guides = "collect")
}

plot_cations("P1") 
plot_cations("P3") 
plot_cations("P5") 
```

## Plot 5: manual temperature data

```{r}
my_pal_2 = brewer.pal(n = 7, "PuBuGn")[5:7]

temp_plot = function(well){
  df1 = manual_well_creek %>% 
    filter(Site == well, Area == "Well") %>% 
    select(Area, Date, Temp_before) %>% 
    drop_na()
  df2 = manual_well_creek %>% 
    filter(Site == well, Area == "Well") %>% 
    select(Area, Date,Temp_after) %>% 
    drop_na()
  df3 = manual_well_creek %>% 
    filter(Site == well, Area == "Creek") %>% 
    select(Area, Date, Temp_before) %>% 
    drop_na()

  ggplot() + 
    geom_point(data = df1, aes(x = Date, y = Temp_before, color = Area, shape = "Before"), size = 2.5) + 
    geom_line(data = df1, aes(x = Date, y = Temp_before, color = Area)) +
    geom_point(data = df2, aes(x = Date, y = Temp_after, color = Area, shape = "After"), size = 2.5) + 
    geom_line(data = df2, aes(x = Date, y = Temp_after, color = Area)) +
    geom_point(data = df3, aes(x = Date, y = Temp_before, color = Area, shape = "Before"), size = 2.5) + 
    geom_line(data = df3, aes(x = Date, y = Temp_before, color = Area)) +
    # scale_color_manual(values = my_pal_2) +
    theme_bw(base_size = 16) +
    scale_x_date(date_breaks = "2 months", date_labels = "%Y-%m") +
    scale_y_continuous(limits = c(0, 25)) +
    # guides(color = "none") +
    labs(title = paste0("Temperature Before & After Pumping at well ", well), x = "Date", y = "Temperature (C)", shape = "Time")
}

p1 = temp_plot("1")
p2 = temp_plot("2")
p3 = temp_plot("3")
p4 = temp_plot("4")
p5 = temp_plot("5")
ggarrange(p1, p2, p3, p4, p5, common.legend = TRUE, legend = "bottom", nrow = 2, ncol = 3)
```

## Plot 6: manual SPC data

```{r}
spc_plot = function(well){
  df1 = manual_well_creek %>% 
    filter(Site == well, Area == "Well") %>% 
    select(Area, Date, SPC_before) %>% 
    drop_na()
  df2 = manual_well_creek %>% 
    filter(Site == well, Area == "Well") %>% 
    select(Area, Date, SPC_after) %>% 
    drop_na()
  df3 = manual_well_creek %>% 
    filter(Site == well, Area == "Creek") %>% 
    select(Area, Date, SPC_before) %>% 
    drop_na()

  ggplot() + 
    geom_point(data = df1, aes(x = Date, y = SPC_before, color = Area, shape = "Before"), size = 2.5) + 
    geom_line(data = df1, aes(x = Date, y = SPC_before, color = Area)) +
    geom_point(data = df2, aes(x = Date, y = SPC_after, color = Area, shape = "After"), size = 2.5) + 
    geom_line(data = df2, aes(x = Date, y = SPC_after, color = Area)) +
    geom_point(data = df3, aes(x = Date, y = SPC_before, color = Area, shape = "Before"), size = 2.5) + 
    geom_line(data = df3, aes(x = Date, y = SPC_before, color = Area)) +
    # scale_color_manual(values = my_pal_2) +
    theme_bw(base_size = 16) +
    scale_x_date(date_breaks = "2 months", date_labels = "%Y-%m") +
    scale_y_continuous(limits = c(400, 900)) +
    # guides(color = "none") +
    labs(title = paste0("SPC Before & After Pumping at well ", well), x = "Date", y = "Specific Conductivity (us/cm)", shape = "Time")
}

p1 = spc_plot("1")
p2 = spc_plot("2")
p3 = spc_plot("3")
p4 = spc_plot("4")
p5 = spc_plot("5")
ggarrange(p1, p2, p3, p4, p5, common.legend = TRUE, legend = "bottom", nrow = 2, ncol = 3)
```